UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-07-03 ms/main.tex
edits throughout

Figure~\ref{fig:umap_ukbb_geo_osgb} presents the UMAP projection from figure~\ref{fig:umap_ukbb_pc10} coloured instead by geographical coordinates from the Ordnance Survey National Grid (OSGB1936), with distances defined as a north or east position relative to the Isles of Scilly. Geographic clusters form in the large White British grouping, reflecting the relationship between genetic and geographic distance, as has been observed in Europe and British-wide data\cite{novembre2008europe,leslie2015fine}. Colouring the projection by distance from London reflects the high migration levels to the city and surrounding area; individuals with more admixed backgrounds tend to have coordinates in the south east of the UK, as visible in figure~\ref{fig:supp_london_distance}. The detailed shape of extended clusters is not stable as we vary the number of PCs included, but the patterns mentioned above are preserved. Figures~\ref{fig:supp_montage_ukbb_eth}, \ref{fig:supp_montage_ukbb_ns}, and~\ref{fig:supp_montage_ukbb_ew} show UMAP plots using the top $40$ PCs from the UKBB.

>>>>

Figure~\ref{fig:umap_ukbb_geo_osgb} presents the UMAP projection from figure~\ref{fig:umap_ukbb_pc10} coloured instead by geographical coordinates from the Ordnance Survey National Grid (OSGB1936), with distances defined as a north or east position relative to the Isles of Scilly. Geographic clusters form in the large White British grouping, reflecting the relationship between genetic and geographic distance, as has been observed in Europe and British-wide data\cite{novembre2008europe,leslie2015fine}. Figure~\ref{fig:umap_ukbb_pc10} shows that the admixed individuals have UMAP coordinates next to White British individuals residing in the South East of the UK, where London is located (see also figure~\ref{fig:supp_london_distance}, where individuals are coloured by distance from London). This likely reflects the high migration levels to the city and surrounding area: the UMAP projection attempted to preserve both the genetic similarity among admixed individuals and the relatedness with White British individuals in cosmopolitan areas.

 The detailed shape of extended clusters is not stable as we vary the number of PCs included, but the patterns mentioned above are preserved. Figures~\ref{fig:supp_montage_ukbb_eth}, \ref{fig:supp_montage_ukbb_ns}, and~\ref{fig:supp_montage_ukbb_ew} show UMAP plots using the top $40$ PCs from the UKBB.
